Refactoring the Candidate Cancer Gene Database

### 2021-05-20 presentation.Rmd
added a few more figs; expanded presentation content

---
title: "Cool Title"
author: "Chris Tastad"
date: "`r format(Sys.time(), '%b %d, %Y')`"
output:
  revealjs::revealjs_presentation:
    center: false
    theme: night
    highlight: zenburn
    transition: none
    reveal_options:
      controls: false
      height: "100%"
      margin: 0.04
      slideNumber: true
      previewLinks: true
---

## Technical Debt fig

## CCGD Overview {.center}

@@

## Application Concept

@@

## Rmarkdown Rendering Flow
<img src="./fig/rmarkdown_flow.svg">

## YAML Configuration Segments

@@

![](./fig/yaml_illustration.svg)

## Subroutine Flow
![](./fig/subroutine_flow.svg)

## Table Reactive Graph
![](./fig/reactive_graph.svg)

## Server-side Processing Flow
![](./fig/parse_flow.svg)

## Search Interface Changes
![](./fig/interface.svg)

## Disk Use Comparison
![](./fig/space_tree.svg)

## Project Directory File Tree Comparison
![](./fig/file_tree.svg)

## Documentation

## app.R code

## markdown code

## List of app improvements

@@

.reveal .progress span {
  background: #7f9f7f;
}

.reveal section img{
  border: 30px solid #e0e1e2;
  border-radius: 10px;
  }

>>>>

---
title: "Refactoring the Candidate Cancer Gene Database"
author: "Christopher Tastad"
date: "May 21, 2021"
output:
  revealjs::revealjs_presentation:
    self_contained: false
    center: false
    theme: night
    highlight: zenburn
    transition: none
    reveal_plugins: ["menu", "notes"]
    reveal_options:
      showNotes: "separate-page"
      controls: true
      height: "100%"
      margin: 0.04
      slideNumber: true
      previewLinks: true
---

## Presentation ToC
> - Original Concept
> - Challenges
> - Refactoring the Concept
> - Outcomes of Refactoring
> - Conclusions

# Original Concept {.center}

--------------------------------------------------------------------------------

--------------------------------------------------------------------------------

## CCGD Overview

--------------------------------------------------------------------------------

@@

--------------------------------------------------------------------------------

## OIT setup

################################################################################

# Challenges {.center}

--------------------------------------------------------------------------------

## Forced Upgrades

--------------------------------------------------------------------------------

## Technical Debt
> - test
> - test

################################################################################

# Refactoring the Concept {.center}

--------------------------------------------------------------------------------

@@

--------------------------------------------------------------------------------

## Front-end Framework
![](./fig/rmarkdown_flow.svg)

#### Rmarkdown rendering flow {.center}

--------------------------------------------------------------------------------

## Front-end Framework

@@

#### YAML configuration segments {.center}

<center>
<img src="./fig/yaml_illustration.svg">
</center>

#### YAML illustration

--------------------------------------------------------------------------------

## Front-end Framework
![](./fig/reactive_graph.svg)

#### table reactive graph {.center}

--------------------------------------------------------------------------------

## Back-end Framework
![](./fig/parse_flow.svg)

#### server-side processing flow {.center}

--------------------------------------------------------------------------------

## Back-end Framework
![](./fig/subroutine_flow.svg)

#### subroutine flow {.center}

################################################################################

# Outcomes of Refactoring {.center}

--------------------------------------------------------------------------------

## Simplified Codebase
```{r echo = T}
df <- read.csv("ccgd_export.csv") # read in base source file
speciesList <- c("Mouse", "Human", "Rat", "Fly", "Fish", "Yeast")

# inputs are arranged in column, width orientation
# each column variable set represents a single input and its params
# next row of inputs layout
ui <- fluidPage( # build shiny app UI
  fluidRow(
    column(
      2,
      selectizeInput("Species",
        label = "Species",
        choices = speciesList
      )
    ),

    column(
      3,
      selectizeInput("Study",
        label = "Study",
        choices = sort(unique(df$Study)),
        selected = NULL,
        multiple = TRUE,
        options = list(placeholder = "All studies")
      )
    ),

    column(
      3,
      selectizeInput("Cancer",
        label = "Cancer",
        choices = sort(unique(df$Cancer)),
        selected = NULL,
        multiple = TRUE,
        options = list(placeholder = "All cancers")
      )
    ),

    column(
      4,
      textAreaInput("Genes",
        label = "Genes",
        placeholder = "GeneA,GeneB,GeneC..."
      )
    )
  ),

  fluidRow(
    column(
      2,
      # button for table export
      downloadButton("downloadData",
        label = "Download"
      )
    )
  ),

  hr(),

  dataTableOutput("searchTable")
)
```

#### app.R Rshiny process snippet {.center}

--------------------------------------------------------------------------------

## Simplified Codebase

```{r echo = T}
---
title: "Candidate Cancer Gene Database"
bibliography: refs/ccgd_paper.bib
nocite: '@*'
output:
    html_document:
        includes:
            in_header: "styles/favicon.html"
            after_body: "styles/footer.html"
        css: styles/styles.css
        theme: readable
---

The Candidate Cancer Gene Database (CCGD) was developed to make accessible a collated set of results from transposon-based forward cancer genetic screens in mice. The value in doing this is to give investigators the ability to quickly filter through the results of many such screens in an effort to determine the candidacy of a gene for its role in cancer.

***

### Background

This work is a product of the [Starr Lab](http://www.med.umn.edu/starrlab/) at the [University of Minnesota](http://www.umn.edu/). The Starr Lab is affiliated with the [Center for Genome Engineering](http://www.cge.umn.edu/), the [Masonic Cancer Center](http://www.cancer.umn.edu/), the Department of [OB/GYN](http://www.obgyn.umn.edu/), and the Department of [Genetics, Cell Biology, and Development](http://www.cbs.umn.edu/gcd/).

##### Contributions

The idea for creating this database originated from Tim Starr, Vincent Keng, and David Largaespada at UMN. Erik Nyre and Ken Abbott developed the original database, with help from Juan Abrahante. A complete website and backend overhaul was conducted by Christopher Tastad in 2019.

* Ken Abbott
* [Erik Nyre](https://www.dorsey.com/people/n/nyre-erik)
* [Juan Abrahante](https://research.umn.edu/units/umii/about-us/staff-directory/juan-e-abrahante-phd)
* [Christopher Tastad](https://github.com/ctastad)

***

##### Updates

The most recent table build was completed on:

<object type="text/html"
        height=40px
        data="build_date.txt"></object>

#### index.html page content snippet {.center}

--------------------------------------------------------------------------------

## Simplified Codebase
![](./fig/file_tree.svg)

#### Project Directory File Tree Comparison {.center}

--------------------------------------------------------------------------------

## Visual Improvements
![](./fig/interface.svg)

#### search interface changes {.center}

--------------------------------------------------------------------------------

## Resource Use Reduction
![](./fig/space_tree.svg)

#### disk use comparison {.center}

--------------------------------------------------------------------------------

## Improved Admin Controls

```{r echo = T}
#   Executed:   locally from the CCGD project dir
#   Options:    -b build (TRUE,FALSE) -s full proj dir sync (TRUE,FALSE)
#               -t table upload file -r reference upload file
#               -k (TURE,FALSE) render site files

# pass arguments from cli
while getopts b:t:r:s: option
do
    case "${option}"
        in
        b) build=${OPTARG};;
        t) table=${OPTARG};;
        r) refs=${OPTARG};;
        s) sync=${OPTARG};;
    esac
done
```

#### ccgd_upload.sh script snippet {.center}

--------------------------------------------------------------------------------

## Sustainable Development
![](./fig/documentation.png)

#### project documentation example {.center}

--------------------------------------------------------------------------------

## Sustainable Development

![](./fig/github.png)

#### public application github repository

--------------------------------------------------------------------------------

## Sustainable Development

![](./fig/release.png)

#### Zenodo-supplied DOI and release for reproducibility {.center}

################################################################################

# Conclusions {.center}

## Conclusions

## Acknowledgements

@@

.reveal section img {
  border: 30px solid #e0e1e2;
  border-radius: 10px;
  }

  .reveal p {
    text-align: left;
  }
  .reveal ul {
    display: block;
  }
  .reveal ol {
    display: block;
  }
